Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
redistributed/removed the initial part of sec 4.5 (before Verification of a controllable system)

The verification of a hybrid continuous/discrete system over an unbounded time, requires the
non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
abstraction of this system by establishing bounds on the error between such a system and a discrete-time/continuous-space model.
This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
the safety specification.
Our second model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a
single iteration, in addition to a stability specification defined by Jury's criteria, and try to synthesise a controller that meets these specs.
A second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time is used to verify that the controller meets the full safety specification. If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

\addtodo{Move as much as possible from Sec~\ref{sec:cof_verification}
to the appendix (e.g. discussion on abstract acceleration). 
Merge the rest in Sec~\ref{sec:cof_synth}.}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{T}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Calculate a set of initial bounds on K based on the input constraints. $$\phi_{input} \Leftrightarrow u_k=r-Kx_k : x \in X^\# \wedge u_k \in [\underline{u}\ \ \overline{u}]$$
\end{enumerate}
\item Next we synthesise a candidate controller. Let $\mathcal{J}\langle I,F \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F \rangle (P_{a-k},\mat{T})\wedge \phi_{input}$ that satisfies the input constraint and Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\phi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \phi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.

>>>>

% The verification of a hybrid continuous/discrete system over an unbounded time, requires the
% non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
% abstraction of this system by establishing bounds on the error between such a system and a discrete-time/continuous-space model.
% This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
% the safety specification.
% The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

% \addtodo{Move as much as possible from Sec~\ref{sec:cof_verification}
% to the appendix (e.g. discussion on abstract acceleration). 
% Merge the rest in Sec~\ref{sec:cof_synth}.}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.

\begin{enumerate}
\item 
We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{T}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Calculate a set of initial bounds on K based on the input constraints. $$\phi_{input} \Leftrightarrow u_k=r-Kx_k : x \in X^\# \wedge u_k \in [\underline{u}\ \ \overline{u}]$$
\end{enumerate}
\item Next we synthesise a candidate controller. 
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a single iteration, in addition to a stability specification defined by Jury's criteria. %, and try to synthesise a controller that meets these specs.

Let $\mathcal{J}\langle I,F \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F \rangle (P_{a-k},\mat{T})\wedge \phi_{input}$ that satisfies the input constraint and Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration. For this purpose, we make use of a second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time. % If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.

\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\phi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \phi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.
